A molecular dynamics study of the solvation of carbon dioxide in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-09-02 ionic_liquids.tex
Revision of the introduction

\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage[version=4]{mhchem}

@@

\section{Introduction}
\ce{CO_2} emissions from fossil-fuel power generation plants account for a large portion of the anthropogenic greenhouse gases \cite{totalenergy}.
International efforts aimed at regulating and reducing \ce{CO_2} emissions are increasing \cite{Tong_2018} and, in this context, the development of technologies for the capture and storage of \ce{CO_2} plays an essential role \cite{Markewitz_2014}.

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures.
The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA).
Despite its high absorption capability, the use of MEA suffers from a serious drawback due to the high cost of solvent recovery \cite{Merkel_2010}.
Other disadvantages include corrosion and degradation by oxidation, solvent loss by evaporation, which also contributes to increase the environmental pollution and generates extra costs for replacing the solvent.

The use of ionic liquids (ILs) as solvents for \ce{CO_2} sequestration occupies a central place in the investigation of alternative technologies.
The reasons are their well known properties, such as their almost negligible vapor pressure and high thermal stability.
Furthermore, they are particularly attractive since they can be designed for specific applications, giving rise to the so-called task-specific ILs \cite{Seo_2014}.
These remarkable properties makes ILs suitable for a wide variety of applications in separation processes \cite{Han_2010,Werner_2010}, catalysis \cite{P_rvulescu_2007}, energy \cite{MacFarlane_2014}, and material science \cite{Mecerreyes_2011,Tom_2015,Dupont_2010,Leones_2017,Kinik_2017}.

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs.
Although most thermodynamic models are formulated in terms of molar fraction, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent.
By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar, or even lower, than that observed for the heavy alkane eicosane.
This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of \ce{[B(CN)_4]^-}.
In this context, molecular simulation provides with a powerful tool for studying solvation processes.
To the best of our knowledge, only one study in literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated.
This motivated us to evaluate different force fields of \ce{[emim][B(CN)_4]} and, as we will show below, it was possible to find a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2}.
For comparison, \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both at infinite dilution and at high concentrations of \ce{CO_2}.
The simulations confirmed what is observed in experiments: for \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}.
The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.

This paper is organized as follows: in Sec.~\ref{sec:theory} we present the theoretical background, which is followed by a brief description of the force fields used in this work.
The methodological details are provided in Sec.~\ref{sec:sim_details} and the corresponding numerical results are given in Sec.~\ref{sec:results}.
Finally, we present some concluding remarks in Sec.~\ref{sec:conclusion}.

>>>>

\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\modulolinenumbers[5]
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage[]{natbib}
\usepackage[version=4]{mhchem}

@@

\section{Introduction}

\ce{CO_2} emissions from fossil-fuel power generation plants account for a large portion of the anthropogenic greenhouse gases \cite{totalenergy}.
International efforts aimed at regulating and reducing \ce{CO_2} emissions are increasing \cite{Tong_2018} and, in this context, the development of technologies for the capture and storage of \ce{CO_2} plays an essential role \cite{Markewitz_2014}.

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures.
The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA).
Despite its high absorption capability, the use of MEA has a serious drawback, which is the high cost of solvent recovery \cite{Merkel_2010}.
Other disadvantages include corrosion and degradation by oxidation, as well as solvent loss by evaporation.
The latter also contributes to the increase of environmental pollution and generates extra costs for replacing the solvent.

The use of ionic liquids (ILs) as solvents for \ce{CO_2} sequestration receives special attention in the investigation of alternative technologies.
The reasons are their well-known properties such as almost negligible vapor pressure and high thermal stability.
Furthermore, they are particularly attractive since they can be designed for specific applications, giving rise to the so-called task-specific ILs \cite{Seo_2014}.
These remarkable properties make ILs suitable for a wide variety of applications in separation processes \cite{Han_2010,Werner_2010}, catalysis \cite{P_rvulescu_2007}, energy management \cite{MacFarlane_2014}, and material science \cite{Mecerreyes_2011,Tom_2015,Dupont_2010,Leones_2017,Kinik_2017}.

Recently, Carvalho \textit{et al}. \cite{Carvalho_2016} called into question the potentiality of ILs for \ce{CO_2} capture.
They warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposedly high absorption capability of certain ILs.
Although most thermodynamic models are formulated in terms of molar fractions, a key factor that dictates the efficiency of absorption process  is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent.
By expressing the solubility values in terms of molality and performing an exhaustive comparison among various ILs, Carvalho \textit{et al}. \cite{Carvalho_2016} show that most systems which promote physical absorption of \ce{CO_2} exhibit solubilities which are similar, or even lower, than that observed for heavy alkanes such as eicosane.
This is true for the popular ILs \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}.
The only exception they found was the family of ILs containing the anion \ce{[B(CN)_4]^-} (tetracyanoborate).

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of \ce{[B(CN)_4]^-}.
In this context, molecular simulation provides us with a powerful tool for studying solvation processes.
To the best of our knowledge, only one study in the literature reports computation of solvation free energy for \ce{CO_2} in \ce{[emim][B(CN)_4]} by molecular simulation \cite{Liu_2014_1}.
Nonetheless, the computed value is overestimated, as observed by deriving a Henry's law constant and comparing it to available experimental data.
Such scarcity motivated us to carry out the present investigation, in which we have evaluated different force fields for \ce{[emim][B(CN)_4]} proposed in the literature.
As we show here, it was possible to find a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2} in this IL.
For comparison, we analyzed the solvation of \ce{CO_2} not only in \ce{[emim][B(CN)_4]}, but also in \ce{[emim][NTf_2]}, both at infinite dilution and at high concentrations.
The simulations confirmed what is observed in experiments.
For \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}, considering a given volume of solvent.
In addition, we also calculated infinite-dilution activity coefficients of other solutes in \ce{[emim][B(CN)_4]} and compared the results with experimental values recently reported in literature.
The solutes we considered were hexane, benzene, cyclohexane, ethanol, and water.

This paper is organized as follows.
In Sec.~\ref{sec:theory}, we present the theoretical background of this work, which is followed by a brief description of the employed force fields in Sec.~\ref{sec:force_field}.
The methodological details are provided in Sec.~\ref{sec:sim_details}, while the corresponding numerical results are given in Sec.~\ref{sec:results}.
Finally, we present some concluding remarks in Sec.~\ref{sec:conclusion}.
